---
bibliography: /home/persican/Documents/library.bib
output:
  github_document:
    html_preview: true
---

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  fig.path = "inst/images/README-"
)
```

```{r setup, echo = FALSE}
library(ggplot2)
theme_set(theme_bw(base_size = 12, base_family = "Open Sans"))
```

## cdom [![Travis-CI Build Status](https://api.travis-ci.org/PMassicotte/cdom.svg?branch=master)](https://travis-ci.org/PMassicotte/cdom) [![AppVeyor Build Status](https://ci.appveyor.com/api/projects/status/github/PMassicotte/cdom?branch=master&svg=true)](https://ci.appveyor.com/project/PMassicotte/cdom) [![Package-License](https://img.shields.io/badge/license-GPL%20%28%3E=%202%29-brightgreen.svg?style=flat)](http://www.gnu.org/licenses/gpl-2.0.html) [![CRAN](http://www.r-pkg.org/badges/version/cdom)](http://cran.rstudio.com/package=cdom) [![Downloads](http://cranlogs.r-pkg.org/badges/cdom?color=brightgreen)](http://www.r-pkg.org/pkg/cdom)


The **cdom** package implements various functions used to model and calculate metrics from absorption spectra of chromophotic dissolved organic matter (CDOM).

This package provides:

1. Simple wrappers to calculate common metrics found in the literature.
    + The **spectral curve** [@Loiselle2009].
    + The **slope ratio (Sr)** [@Helms2008].
    + The **spectral slope (S)** [@Jerlov1968; @Lundgren1976; @Bricaud1981].

2. The function to use the **Gaussian decomposition approach** proposed in Massicotte and Markager, (2015).

The package can be installed using the following command.

```{r install, eval = FALSE}
devtools::install_github("PMassicotte/cdom")
```

Please note that this is a developing version of the package for testing only. Please fill an issue when you find bugs.

All functions from the package start with the `cdom_` prefix.

```{r}
library(cdom)
ls("package:cdom")
```

# Examples

## Gaussian decomposition

**This should be used with care since it is in very early developpement.**

 
```{r}
library(dplyr)
library(gsubfn)

set.seed(1234)

wl <- seq(190, 700, length.out = 200)

# Build the base exp model
a0  <- 10
S <- 0.02
K <- 0

exp_part <- a0 * exp(-S * (wl - 350)) + K
exp_part <- exp_part + rnorm(wl, sd = 0.001)

# Build 3 Gaussian components
p0 <- c(43, 60, 70)
p1 <- c(250, 350, 400)
p2 <- c(10, 22, 15)

params <- c(a0, S, K, as.vector(rbind(p0, p1, p2)))
ngaussian <- (length(params) - 3) / 3
n <- c("a0", "S", "K", paste0(rep(c("p0", "p1", "p2"), time = ngaussian),
       rep(letters[1:ngaussian], each = 3)))
params <- setNames(params, n)

myfunc <- build_model(ngaussian = ngaussian)

fo <- formula(sub(".*~", "~", deparse(myfunc)))
func <- gsubfn::fn$identity(fo)
spc <- do.call(func, c(list(x = wl), params))

# Plot a preview

plot(wl, spc, type = "l")

# Decompose the generated spectra

myfit <- cdom_gaussian(x = wl, y = spc, min_distance = 50)

coef(myfit)

plot(myfit)
```

## The spectral slope (S)

The `cdom_fit_exponential()` function fits an exponential curve to CDOM data using the simple model proposed by @Jerlov1968, @Lundgren1976, @Bricaud1981.

```tex
a(\lambda) = a(\lambda0)e^{-S(\lambda - \lambda0)} + K
```

```{r exponential}
library(ggplot2)
library(cdom)
data("spectra")

fit <- cdom_fit_exponential(wl = spectra$wavelength,
                       absorbance = spectra$spc3,
                       wl0 = 350,
                       startwl = 190,
                       endwl = 900)

ggplot(spectra, aes(x = wavelength, y = spc3)) +
  geom_point() +
  geom_line(aes(y = fit$data$.fitted), col = "red") +
  xlab("Wavelength (nm)") +
  ylab(expression(paste("Absorption (", m ^ {-1}, ")")))

```

## The slope ratio (SR)

The `cdom_slope_ratio()` function calculates the slope ratio (S<sub>R</sub>) which is defined as: S<sub>275-295</sub>/S<sub>350-400</sub>. See @Helms2008 for detailed information.

```{r sr}
library(cdom)
data("spectra")

cdom_slope_ratio(spectra$wavelength, spectra$spc1)
```

## The spectral curve

The `cdom_spectral_curve()` function generates the spectral curve using the slope of the linear regression between the natural log absorption spectrum and wavelengths over a sliding window of 21 nm interval (default) at 1 nm resolution. See @Loiselle2009 for detailed information.

```{r spectral_curve}
library(cdom)
data("spectra")

res <-  cdom_spectral_curve(wl = spectra$wavelength,
                       absorbance = spectra$spc10,
                       interval = 21,
                       r2threshold = 0.98) # Maybe to restrictive...

ggplot(res, aes(x = wl, y = s)) +
  geom_point() +
  geom_line() +
  xlab("Wavelength (nm)") +
  ylab(expression(paste("Spectral slope (", nm ^ {-1}, ")")))

```

# Data

A total 25 absorption spectra are provided in the package.

```{r data}
library(ggplot2)
library(tidyr)
data("spectra")

spectra <- gather(spectra, sample, absorption, -wavelength)

ggplot(spectra, aes(x = wavelength, y = absorption, group = sample)) +
  geom_line(size = 0.1) +
  xlab("Wavelength (nm)") +
  ylab(expression(paste("Absorption (", m ^ {-1}, ")")))
```

# How to cite the package

```{r}
citation("cdom")
```


# References
